---
title: "PA Design Reference Manual - Project Status & Tracking"
author: "Project Status Dashboard"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    theme: cosmo
    highlight: tango
    code_folding: hide
    df_print: paged
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(knitr)
```

---

# 📊 Project Overview

## Project Information
- **Project Name**: PA Design Reference Manual
- **Project Start Date**: `r Sys.Date()`
- **Target Completion**: 3 months (May 1, 2026)
- **Status**: ✅ APPROVED - Phase 1 Initiation
- **Current Phase**: Setup & Planning
- **Next Milestone**: Data Extraction from IFX_2022_2025/02_Projects

---

## 🎯 Project Objectives

### Primary Objectives
1. **Reference Manual**: Create comprehensive PA design notes from actual work experience
2. **Interview Preparation**: Structured knowledge base combining textbook + research + practical experience
3. **Automation Framework**: Develop fool-proof prompt/plan for automated PA design apps

### Guiding Principles ⭐
- **Quality**: No compromises on accuracy and depth
- **Truthfulness**: Fact-based, evidence-driven content
- **Rationale**: Clear reasoning behind every design decision
- **Simplicity**: Complex concepts explained clearly
- **Professional Workflow**: Industry-standard practices
- **Easy to Use**: Intuitive navigation and structure
- **Flexibility**: Adaptable to different use cases
- **Incremental Steps**: Small, verifiable progress

### Technical Requirements 🔧
- **Platform Independence**: Works on any system/PC
- **Reusable Core Logic**: Modular, library-based architecture
- **Scalable Design**: Extensible for future enhancements
- **Developer-Friendly**: Clear documentation for handover

---

# 📅 Project Timeline

## Overall Schedule
```{r timeline}
timeline <- data.frame(
  Phase = c("Planning & Setup", "Manual Development", "Automation Framework", "Testing & Refinement"),
  Duration = c("Week 1-2", "Week 3-8", "Week 9-11", "Week 12"),
  Status = c("🟢 In Progress", "⚪ Not Started", "⚪ Not Started", "⚪ Not Started"),
  Completion = c("20%", "0%", "0%", "0%")
)
kable(timeline, caption = "Project Phase Timeline")
```

## Milestone Schedule
- **Week 1-2**: Project setup, data extraction, template creation *(Current)*
- **Week 3-4**: Chapter 1-2 development (Fundamentals, Load-Pull)
- **Week 5-6**: Chapter 3-4 development (Linearization, Efficiency)
- **Week 7-8**: Chapter 5-6 development (Thermal, Integration)
- **Week 9-11**: Automation framework development
- **Week 12**: Final review and refinement

## Check-in Schedule
- **Frequency**: Every 4 days
- **Duration**: ~1 hour per check-in
- **Weekly Commitment**: 4 hours
- **Next Check-in**: `r Sys.Date() + 4` days

---

# ✅ Completed Activities

## Session 1: `r Sys.Date()`

### What Was Completed
1. ✅ **Folder Structure Analysis**
   - Explored IFX_2022_2025/02_Projects folder
   - Identified Tx_Baseline project structure
   - Located key documents and simulation files

2. ✅ **Reference Analysis**
   - Analyzed Atomic_Cosmic_RFView.Rmd structure
   - Identified successful patterns: tabsets, interactive navigation, self-contained HTML
   - Documented styling and organization approach

3. ✅ **Project Plan Creation**
   - Created comprehensive 6-chapter manual structure
   - Designed 4-phase automation framework
   - Developed execution roadmap with milestones

4. ✅ **Stakeholder Approval**
   - Plan reviewed and approved
   - Timeline agreed: 3 months
   - Check-in frequency: Every 4 days
   - Approach confirmed: Option A (Manual first, automation later)

### Files Created
- `/PA_Design_Reference_Manual/` - Project folder
- `PA_Design_Project_Plan.Rmd` - Comprehensive project plan
- `PA_Design_Project_Plan.html` - Rendered plan document
- `PROJECT_STATUS.Rmd` - This tracking document (in progress)

### Data Sources Identified
```{r data_sources}
data_sources <- data.frame(
  Source = c(
    "IFX_2022_2025/02_Projects/01_Tx_Baseline/",
    "IFX_2022_2025/00_Master_html_file/",
    "IFX_2022_2025/07_Technical_reports/",
    "IFX_2022_2025/03_PRD/"
  ),
  Type = c(
    "Project files, simulations",
    "Compiled HTML reports",
    "Technical documentation",
    "Internal guidelines"
  ),
  Status = c("✅ Identified", "✅ Identified", "✅ Identified", "✅ Identified")
)
kable(data_sources, caption = "Data Sources for Manual Development")
```

### Key Decisions Made
1. **Structure**: 6-chapter manual (Fundamentals → Load-Pull → Linearization → Efficiency → Thermal → Integration)
2. **Approach**: Sequential development (manual first, automation second)
3. **Format**: Self-contained HTML with tabsets, similar to Atomic_Cosmic_RFView.Rmd
4. **Quality Framework**: Evidence-based, with rationale for every design choice
5. **Tracking**: Dedicated status document + developer handover documentation

---

# 🔄 Current Status

## Active Tasks (Week 1)
```{r current_tasks}
current <- data.frame(
  Task = c(
    "Create PROJECT_STATUS.Rmd",
    "Create DEVELOPER_GUIDE.md",
    "Set up project folder structure",
    "Extract data from Tx_Baseline project",
    "Create Chapter 1 template"
  ),
  Owner = c("System", "System", "System", "System", "System"),
  Status = c("🟡 In Progress", "⚪ Not Started", "⚪ Not Started", "⚪ Not Started", "⚪ Not Started"),
  Priority = c("High", "High", "High", "High", "Medium"),
  Due = c("Today", "Today", "Feb 2", "Feb 3-4", "Feb 5")
)
kable(current, caption = "Current Sprint Tasks")
```

## Blockers & Risks
- **None identified yet**

## Waiting On
- **None** - Ready to proceed

---

# 🎯 Next Steps

## Immediate Actions (Next 4 Days)
1. **Complete Setup**
   - ✅ Finalize PROJECT_STATUS.Rmd
   - 📝 Create DEVELOPER_GUIDE.md
   - 📝 Set up modular folder structure

2. **Data Extraction Phase**
   - 📝 Extract Tx_Baseline project information
   - 📝 Catalog simulation files and results
   - 📝 Document measurement data
   - 📝 Extract design decisions and rationale

3. **Template Creation**
   - 📝 Create Chapter 1 Rmd template
   - 📝 Define reusable CSS styles
   - 📝 Set up bibliography system
   - 📝 Create figure/table numbering system

## Upcoming Milestones (Week 2)
- Begin Chapter 1: PA Fundamentals
- Extract theoretical content
- Integrate practical evidence from projects
- Create first draft for review

---

# 📊 Progress Metrics

## Overall Completion
```{r progress}
progress_data <- data.frame(
  Component = c("Planning", "Manual Development", "Automation", "Testing"),
  Target = c(100, 0, 0, 0),
  Actual = c(90, 0, 0, 0),
  Remaining = c(10, 100, 100, 100)
)
kable(progress_data, caption = "Progress by Component (%)")
```

**Overall Project Completion**: 7.5% (Planning phase nearly complete)

## Quality Metrics
- **Accuracy**: Target 100% (fact-checked against source data)
- **Completeness**: Target 100% (all identified topics covered)
- **Usability**: Target 95% (intuitive navigation, clear explanations)
- **Reusability**: Target 90% (modular code, clear documentation)

---

# 💡 Highlights & Insights

## Wins 🎉
1. **Clear Vision**: Well-defined objectives and guiding principles
2. **Structured Approach**: Systematic plan approved by stakeholder
3. **Rich Data Source**: Comprehensive project files available
4. **Proven Template**: Successful Atomic_Cosmic_RFView.Rmd as reference

## Challenges 🤔
1. **Data Volume**: Large amount of information to process systematically
2. **Quality vs Speed**: Balancing thoroughness with timeline
3. **Automation Complexity**: Need to design fool-proof prompting system

## Lessons Learned 📚
- Early stakeholder alignment critical
- Need robust tracking from day one
- Incremental approach reduces risk
- Platform independence requires careful architecture planning

---

# 📁 File & Resource Tracker

## Project Files Created
```{r files_created}
files <- data.frame(
  File = c(
    "PA_Design_Project_Plan.Rmd",
    "PA_Design_Project_Plan.html",
    "PROJECT_STATUS.Rmd",
    "PROJECT_STATUS.html (pending)"
  ),
  Purpose = c(
    "Master project plan document",
    "Rendered plan for review",
    "Live project tracking",
    "Status dashboard"
  ),
  Status = c("✅ Complete", "✅ Complete", "🟡 In Progress", "⚪ Pending"),
  Size = c("~15 KB", "~200 KB", "~10 KB", "TBD")
)
kable(files, caption = "Project Documentation Files")
```

## Data Files Used
- **IFX_2022_2025/02_Projects/01_Tx_Baseline/** - Primary data source
  - Planning documents
  - Simulation files (ADS, SystemVue)
  - Measurement results
  - Design documentation

## External Resources
- Atomic_Cosmic_RFView.Rmd - Template reference
- styles.css - Styling reference

---

# 🔧 Technical Stack & Dependencies

## Current Tools
- **R Markdown**: Document creation
- **Knitr**: Document rendering
- **RStudio/VS Code**: Development environment
- **Git**: Version control

## Required Packages
```{r dependencies, echo=TRUE, eval=FALSE}
# R packages needed
install.packages(c(
  "rmarkdown",    # Document rendering
  "knitr",        # Report generation
  "htmltools",    # HTML manipulation
  "DT",           # Interactive tables
  "plotly",       # Interactive plots
  "ggplot2",      # Static plots
  "kableExtra"    # Enhanced tables
))
```

## Platform Independence
- Self-contained HTML output
- No external dependencies for viewing
- Standard R packages for development
- Works on Windows, macOS, Linux

---

# 📝 Action Items for Next Check-in

## By `r Sys.Date() + 4`
- [ ] Complete DEVELOPER_GUIDE.md
- [ ] Extract all Tx_Baseline data
- [ ] Create Chapter 1 template
- [ ] Begin Chapter 1 content development
- [ ] Update PROJECT_STATUS with progress

## Review Questions
1. Is the data extraction capturing the right level of detail?
2. Does the Chapter 1 template meet expectations?
3. Any changes to structure or approach needed?

---

# 📧 Contact & Updates

## How to Update This Document
1. Open `PROJECT_STATUS.Rmd` in RStudio/VS Code
2. Update relevant sections
3. Render: `rmarkdown::render("PROJECT_STATUS.Rmd")`
4. Commit changes to Git

## Quick Status Check
```{r quick_status, echo=TRUE, eval=FALSE}
# Run this in R console for quick status
cat(paste0(
  "Project: PA Design Reference Manual\n",
  "Phase: Setup & Planning\n",
  "Completion: 7.5%\n",
  "Next Milestone: Data Extraction (Feb 3-4)\n",
  "Status: ✅ On Track\n"
))
```

---

*Last Updated: `r Sys.Date()` | Auto-generated from PROJECT_STATUS.Rmd*
