---
title: "PA Design Reference Manual - Project Status & Tracking"
author: "Project Status Dashboard"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    theme: cosmo
    highlight: tango
    code_folding: hide
    df_print: paged
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(knitr)
```

---

# 📊 Project Overview

## Project Information
- **Project Name**: PA Design Reference Manual
- **Project Start Date**: `r Sys.Date()`
- **Target Completion**: 3 months (May 1, 2026)
- **Status**: ✅ APPROVED - Phase 1 Initiation
- **Current Phase**: Setup & Planning
- **Next Milestone**: Data Extraction from IFX_2022_2025/02_Projects

---

## 🎯 Project Objectives

### Primary Objectives
1. **Reference Manual**: Create comprehensive PA design notes from actual work experience
2. **Interview Preparation**: Structured knowledge base combining textbook + research + practical experience
3. **Automation Framework**: Develop fool-proof prompt/plan for automated PA design apps

### Guiding Principles ⭐
- **Quality**: No compromises on accuracy and depth
- **Truthfulness**: Fact-based, evidence-driven content
- **Rationale**: Clear reasoning behind every design decision
- **Simplicity**: Complex concepts explained clearly
- **Professional Workflow**: Industry-standard practices
- **Easy to Use**: Intuitive navigation and structure
- **Flexibility**: Adaptable to different use cases
- **Incremental Steps**: Small, verifiable progress

### Technical Requirements 🔧
- **Platform Independence**: Works on any system/PC
- **Reusable Core Logic**: Modular, library-based architecture
- **Scalable Design**: Extensible for future enhancements
- **Developer-Friendly**: Clear documentation for handover

---

# 📅 Project Timeline

## Overall Schedule
```{r timeline}
timeline <- data.frame(
  Phase = c("Planning & Setup", "Manual Development", "Automation Framework", "Testing & Refinement"),
  Duration = c("Week 1-2", "Week 3-8", "Week 9-11", "Week 12"),
  Status = c("🟢 Complete", "🟡 In Progress", "⚪ Not Started", "⚪ Not Started"),
  Completion = c("75%", "5%", "0%", "0%")
)
kable(timeline, caption = "Project Phase Timeline")
```

## Milestone Schedule
- **Week 1-2**: Project setup, data extraction plans, chapter templates *(COMPLETE - 75%)*
- **Week 3-4**: Chapter 1-2 data extraction and development (Fundamentals, Load-Pull) *(NEXT)*
- **Week 5-6**: Chapter 3-4 development (Linearization, Efficiency)
- **Week 7-8**: Chapter 5-6 development (Thermal, Integration)
- **Week 9-11**: Automation framework development
- **Week 12**: Final review and refinement

## Check-in Schedule
- **Frequency**: Every 4 days
- **Duration**: ~1 hour per check-in
- **Weekly Commitment**: 4 hours
- **Next Check-in**: `r Sys.Date() + 4` days

---

# ✅ Completed Activities

## Session 1: `r Sys.Date()`

### What Was Completed
1. ✅ **Folder Structure Analysis**
   - Explored IFX_2022_2025/02_Projects folder
   - Identified two core projects:
     * 01_Tx_Baseline: GaN transistor characterization (7 sub-folders, 30+ PDFs)
     * 02_PAM_B: Power amplifier module design (10 sub-folders, 30+ PDFs)
   - Mapped project relationship: Tx_Baseline feeds into PAM_B

2. ✅ **Comprehensive Chapter Structure**
   - Created 6-chapter structure based on actual project data
   - **Part I: Transistor Design** (Chapters 1-2)
     * Chapter 1: GaN Fundamentals, Modeling, Downselection, Matching
     * Chapter 2: Advanced topics (Thermal, Stability, Reliability)
   - **Part II: PA Module Design** (Chapters 3-4)
     * Chapter 3: Architecture, Multi-stage design, Layout/EM, Assembly, Tuning/CV
     * Chapter 4: Advanced PA topics (DPD, ET, Doherty, mmWave)
   - **Part III: Automation** (Chapters 5-6)
     * Chapter 5: Automation principles, optimization, template/guided/autonomous design
     * Chapter 6: Future trends and career development
   - Mapped data sources to each section
   - Defined learning objectives for each subsection

3. ✅ **Data Source Identification**
   - **Tx_Baseline Project** (01_Tx_Baseline/):
     * 01_Background_overview/ - 4 overview PDFs
     * 02_Design/ - 17 DOE builds with simulations
     * 03_Measurements/ - Downselection reports
     * 04_Model/ - Device modeling data
     * 05_Assembly/ - Assembly procedures
     * 06_Minipac_V02/ - Package documentation
     * 07_PCB-pac/ - Baseline results
   - **PAM_B Project** (02_PAM_B/):
     * 01_Overview/ - Flow charts and summary
     * 02_Mini-pac_PAM_B/ - Package details
     * 03_Design/ - System-level design
     * 04_Layout/ - Layout files
     * 05_Tapeout/ - EM simulations, final layouts
     * 06_Assembly/ - Assembly procedures
     * 07_Tuning_CV/ - Comprehensive tuning data (BOM variants, CV execution)
     * 08_Results_Analysis/ - Performance analysis
     * 09_Sim_vs_Meas/ - Correlation analysis
     * 10_Poster/ - Summary presentation

4. ✅ **Document Structure**
   - Created CHAPTER_STRUCTURE.Rmd with complete outline
   - Rendered as HTML for easy review
   - Each chapter section includes:
     * Learning objectives
     * Content structure (concept → theory → practice → evidence)
     * Data source paths
     * Figures to generate
     * Factual evidence requirements

### Key Insights from Data Analysis

#### Transistor Design (Tx_Baseline)
- **DOE Approach**: Systematic exploration through 17 builds (DOE1-DOE17)
- **Device Sizes**: Range from 2.4mm to 12mm gate periphery
- **Technologies**: GaN HEMT variants (T9095A, T9507B, R9505, T9504, etc.)
- **Downselection**: Structured methodology with clear criteria
- **Design Flow**: Specifications → Simulations → Measurements → Downselection

#### PA Module Design (PAM_B)
- **Architecture**: Multi-stage (driver + final stage)
- **Variants**: Multiple BOM variants (NIJ, J01F, J05, J07, J08, J09)
- **Process Flow**: Design → Layout → Tapeout → Assembly → Tuning → CV → Analysis
- **EM Analysis**: Comprehensive studies (dielectric thickness, molding material impacts)
- **Tuning Process**: Well-documented CV execution with BOM tracking
- **Validation**: Simulation vs measurement correlation

#### Automation Opportunities Identified
1. **Template-based**: DOE structure is highly repeatable
2. **Parametric Sweeps**: Component value optimization
3. **EM Simulation**: Automated variant generation
4. **Data Analysis**: Tuning data crunching and visualization

5. ✅ **Complete Document Inventory (Steps 9/10 & 10/10)**
   - Created comprehensive inventory of ALL documents in both projects
   - **Tx_Baseline**: Catalogued 50-60 primary documents
     * Background (4 docs) → Design (20+ docs) → Measurements (15+ docs)
     * Modeling (7+ docs) → Assembly (multiple)
   - **PAM_B**: Catalogued 80-100 primary documents
     * Overview → Design (6 categories) → Tapeout (7+ docs)
     * Tuning/CV (20+ docs) → Results (50+ analysis docs)
   - **Total**: 130-160 technical documents inventoried
   - Created structured reading order with priority levels (High/Medium/Low)
   - Defined data extraction template for systematic content capture
   - Document location: `DOCUMENT_INVENTORY.md`

### Documents Created This Session
1. **PROJECT_STATUS.Rmd** - This tracking document (460+ lines)
2. **DEVELOPER_HANDOVER.md** - Complete developer documentation
3. **CHAPTER_STRUCTURE.Rmd** - Detailed 6-chapter outline (600+ lines)
4. **DOCUMENT_INVENTORY.md** - Complete file inventory (800+ lines)
5. **chapter_template.Rmd** - Reusable chapter template
6. **verify_structure.R** - Automated verification script
7. **AUTOMATION_FRAMEWORK_PROMPTS.md** - Enhanced automation framework (850 lines)
8. **linearity_optimizer.py** - Python optimization tool (850 lines)
9. **plot_tradeoffs.R** - R visualization tool (550 lines)
10. **ENHANCEMENT_SUMMARY.md** - Enhancement documentation

### Folder Structure Created
```
PA_Design_Reference_Manual/
├── PROJECT_STATUS.Rmd (this file)
├── DEVELOPER_HANDOVER.md
├── DOCUMENT_INVENTORY.md
├── CHAPTER_STRUCTURE.Rmd
├── CHAPTER_STRUCTURE.html
├── AUTOMATION_FRAMEWORK_PROMPTS.md ← NEW
├── ENHANCEMENT_SUMMARY.md ← NEW
├── Scripts/
│   ├── linearity_optimizer.py ← NEW
│   └── plot_tradeoffs.R ← NEW
├── chapter_template.Rmd
├── verify_structure.R
├── Chapters/
│   ├── Chapter_01/
│   ├── Chapter_02/
│   ├── Chapter_03/
│   ├── Chapter_04/
│   ├── Chapter_05/
│   └── Chapter_06/
├── Output/
└── Supporting_docs/
```

### Data Sources Mapped
```{r data_sources}
data_sources <- data.frame(
  Source = c(
    "IFX_2022_2025/02_Projects/01_Tx_Baseline/",
    "IFX_2022_2025/00_Master_html_file/",
    "IFX_2022_2025/07_Technical_reports/",
    "IFX_2022_2025/03_PRD/"
  ),
  Type = c(
    "Project files, simulations",
    "Compiled HTML reports",
    "Technical documentation",
    "Internal guidelines"
  ),
  Status = c("✅ Identified", "✅ Identified", "✅ Identified", "✅ Identified")
)
kable(data_sources, caption = "Data Sources for Manual Development")
```

### Key Decisions Made
1. **Structure**: 6-chapter manual (Fundamentals → Load-Pull → Linearization → Efficiency → Thermal → Integration)
2. **Approach**: Sequential development (manual first, automation second)
3. **Format**: Self-contained HTML with tabsets, similar to Atomic_Cosmic_RFView.Rmd
4. **Quality Framework**: Evidence-based, with rationale for every design choice
5. **Tracking**: Dedicated status document + developer handover documentation

---

# 🔄 Current Status

## Active Tasks (Week 2)
```{r current_tasks}
current <- data.frame(
  Task = c(
    "Explore project folder structures",
    "Create comprehensive chapter structure",
    "Create Tx_Baseline extraction plan",
    "Create PAM_B extraction plan",
    "Create chapter 1 template",
    "Create chapters 2-6 outlines",
    "Reorganize chapter 3 structure",
    "Begin manual PDF extraction"
  ),
  Owner = c("System", "System", "System", "System", "System", "System", "System", "User"),
  Status = c("✅ Complete", "✅ Complete", "✅ Complete", "✅ Complete", "✅ Complete", "✅ Complete", "✅ Complete", "🟡 In Progress"),
  Priority = c("High", "High", "High", "High", "High", "High", "Critical", "High"),
  Due = c("Done", "Done", "Done", "Done", "Done", "Done", "Done", "Week 2")
)
kable(current, caption = "Current Sprint Tasks")
```

## Phase 1 Setup - COMPLETE ✅
1. ✅ **Planning & Structure**
   - Approved comprehensive plan with 6-chapter manual + automation framework
   - Created detailed chapter structure with learning objectives
   - Established project tracking and developer handover system

2. ✅ **Data Discovery**
   - Explored both projects (Tx_Baseline & PAM_B)
   - Identified 130-160 technical documents
   - Created systematic extraction plans:
     * **Tx_Baseline**: 10 extraction files (80KB), 6 DOE templates, progress tracking
     * **PAM_B**: Comprehensive 5-week extraction plan (50-75 hours), 16 priority folders mapped

3. ✅ **Chapter Templates & Outlines**
   - **Chapter 1**: Complete template (1,157 lines, 72KB) ready for Tx_Baseline data
   - **Chapters 2-6**: Detailed outlines (~63KB total) with:
     * Content structure and data source mappings
     * Figure specifications and table templates
     * Cross-reference integration points
     * Learning objectives for each section
   
4. ✅ **Critical Reorganization**
   - **Design Verification Before Fabrication**: Moved linearity, stability, reliability analysis from post-fabrication (Chapter 4) to design phase (Chapter 3)
   - **PAM_B Folder Alignment**: Restructured Chapter 3 to match actual project workflow (03_Design/ folder structure)
   - **Updated PA_Design_Project_Plan.Rmd**: Rendered with corrected chapter organization
   - Created complete document inventory with priority levels
   - Mapped data sources to chapter sections

3. ✅ **Infrastructure**
   - Created folder structure (6 chapter folders + outputs)
   - Set up templates and verification scripts
   - Established documentation standards

## Key Findings
1. **Tx_Baseline**: 17 DOE builds provide rich dataset for transistor design chapter
2. **PAM_B**: Complete design-to-validation flow documented in 10 sub-folders
3. **Relationship**: Transistor downselection (Tx_Baseline) directly feeds PA module design (PAM_B)
4. **Automation Ready**: Repetitive processes identified (DOE, EM sims, tuning optimization)

## Blockers & Risks
- **None identified** - All data sources accessible
- Phase 1 setup complete with no blockers

## Waiting On
- **User approval** - Review of rendered documents before proceeding to content extraction

---

# 🎯 Next Steps

## Phase 2: Content Extraction (Starting Next Session)

### Immediate Actions (Next Check-in: Feb 5, 2026)

1. **HIGH PRIORITY Documents - Read & Extract** *(Week 2)*
   - Tx_Baseline: `01_Background_overview/*.pdf` (4 documents)
   - Tx_Baseline: `03_Measurements/03_Down_selection/10_Main_selection.PNG`
   - PAM_B: `01_Overview/` and `02_Mini-pac_PAM_B/01_Introduction/`
   - PAM_B: `08_Results_Analysis/01_Summary.pdf`
   - **Deliverable**: Foundation content for Chapters 1 & 3

2. **Chapter 1 Initial Draft** *(Week 2-3)*
   - Section 1.1: Introduction & Context (from background docs)
   - Section 1.2: GaN Device Fundamentals (extract from technical reports)
   - Section 1.3: Device Characterization (from measurement procedures)
   - **Deliverable**: Chapter 1 first draft (30-40% complete)

3. **Chapter 3 Initial Draft** *(Week 2-3)*
   - Section 3.1: PA Architecture Overview (from PAM_B overview)
   - Section 3.2: Multi-stage Design Approach (from design docs)
   - **Deliverable**: Chapter 3 first draft (30-40% complete)

## Week 2 Goals
- Complete document inventory for both projects
- Extract and organize all specifications
- Begin content development for Chapters 1 & 3
- Create first set of figures and tables
- Prepare initial draft sections for review

## Check-in #1 Agenda (4 days)
1. Review document inventory completeness
2. Discuss extracted specifications and requirements
3. Review Chapter 1 & 3 initial content
4. Address any ambiguities in source documents
5. Adjust plan if needed based on findings

---

# 📊 Progress Metrics

## Overall Completion
```{r progress}
progress_data <- data.frame(
  Component = c("Planning & Setup", "Data Extraction", "Chapter Development", "Automation Framework", "Review & Refinement"),
  Target_Hours = c(16, 24, 80, 40, 16),
  Actual_Hours = c(12, 4, 4, 0, 0),
  Percent_Complete = c(75, 15, 5, 0, 0)
)
kable(progress_data, caption = "Progress by Component")
```

**Overall Project Completion**: ~15% (Week 2 - Planning complete, extraction infrastructure ready)

**Estimated Total Effort**: 176 hours over 3 months (~15 hours/week average)

## Quality Metrics
- **Accuracy**: Target 100% (fact-checked against source data)
- **Completeness**: Target 100% (all identified topics covered)
- **Usability**: Target 95% (intuitive navigation, clear explanations)
- **Reusability**: Target 90% (modular code, clear documentation)

---

# 💡 Highlights & Insights

## Wins 🎉
1. **Comprehensive Structure**: 6-chapter outline with clear learning path
2. **Rich Data Sources**: 80+ documents in Tx_Baseline, 80-100 in PAM_B identified
3. **Clear Relationships**: Transistor → PA module design flow understood
4. **Automation Roadmap**: Template → Guided → Autonomous phases defined
5. **Stakeholder Alignment**: Plan approved with clear objectives
6. **✨ NEW: Extraction Infrastructure**: 
   - Tx_Baseline plan (10 files, 80KB) + PAM_B plan (5-week roadmap)
   - Chapter 1 template complete (1,157 lines)
   - Chapters 2-6 outlines complete (~63KB)
7. **✨ NEW: Critical Reorganization**: 
   - Design verification correctly placed BEFORE fabrication
   - Chapter structure aligned with actual PAM_B workflow
   - Folder 03_Design/ structure now drives Chapter 3 organization

## Challenges 🤔
1. **Data Volume**: 60+ PDFs to process systematically
2. **Quality vs Speed**: Maintaining accuracy while meeting 3-month timeline
3. **Technical Depth**: Balancing expert-level detail with accessibility
4. **Automation Complexity**: Designing fool-proof prompts for diverse use cases

## Lessons Learned 📚
- **Systematic Exploration**: Folder structure reveals design methodology
- **Documentation Value**: Well-organized projects enable knowledge extraction
- **Dual Goals**: Manual development and automation framework are complementary
- **Incremental Progress**: Small, verifiable steps prevent overwhelming scope
- Need robust tracking from day one
- Incremental approach reduces risk
- Platform independence requires careful architecture planning

---

# 📁 File & Resource Tracker

## Project Files Created
```{r files_created}
files <- data.frame(
  File = c(
    "PA_Design_Project_Plan.Rmd",
    "PA_Design_Project_Plan.html",
    "PROJECT_STATUS.Rmd",
    "PROJECT_STATUS.html (pending)"
  ),
  Purpose = c(
    "Master project plan document",
    "Rendered plan for review",
    "Live project tracking",
    "Status dashboard"
  ),
  Status = c("✅ Complete", "✅ Complete", "🟡 In Progress", "⚪ Pending"),
  Size = c("~15 KB", "~200 KB", "~10 KB", "TBD")
)
kable(files, caption = "Project Documentation Files")
```

## Data Files Used
- **IFX_2022_2025/02_Projects/01_Tx_Baseline/** - Primary data source
  - Planning documents
  - Simulation files (ADS, SystemVue)
  - Measurement results
  - Design documentation

## External Resources
- Atomic_Cosmic_RFView.Rmd - Template reference
- styles.css - Styling reference

---

# 🔧 Technical Stack & Dependencies

## Current Tools
- **R Markdown**: Document creation
- **Knitr**: Document rendering
- **RStudio/VS Code**: Development environment
- **Git**: Version control

## Required Packages
```{r dependencies, echo=TRUE, eval=FALSE}
# R packages needed
install.packages(c(
  "rmarkdown",    # Document rendering
  "knitr",        # Report generation
  "htmltools",    # HTML manipulation
  "DT",           # Interactive tables
  "plotly",       # Interactive plots
  "ggplot2",      # Static plots
  "kableExtra"    # Enhanced tables
))
```

## Platform Independence
- Self-contained HTML output
- No external dependencies for viewing
- Standard R packages for development
- Works on Windows, macOS, Linux

---

# 🎉 Recent Enhancements (February 1, 2026)

## Automation Framework Enhancement - COMPLETED ✅

### What Was Enhanced
User requested enhancements to automation framework focusing on three critical areas:

1. **Linearity (Very Important)**
   - ✅ Added comprehensive IM3/ACPR/EVM specifications
   - ✅ Created linearity-focused design prompts
   - ✅ Bias point optimization for linearity vs efficiency
   - ✅ Two-tone and modulated signal analysis procedures
   - ✅ DPD capability assessment framework

2. **Trade-off Plots for Better Design Decisions**
   - ✅ Performance Triangle visualization (Efficiency-Linearity-Power)
   - ✅ Pareto front analysis for multi-objective optimization
   - ✅ PAE vs IM3 trade-off plots
   - ✅ Pout vs Linearity with bias coloring
   - ✅ Design space contour maps (IM3, PAE vs Iq, ZL)
   - ✅ Yield vs Performance Margin curves

3. **Mass Manufacturing Perspective**
   - ✅ Yield prediction (Monte Carlo with 10k samples)
   - ✅ Process capability analysis (Cp/Cpk)
   - ✅ Cost analysis at different volumes (1k, 10k, 100k)
   - ✅ Design for Manufacturing (DFM) checklists
   - ✅ Tuning vs tuning-free decision trees
   - ✅ Reliability/MTBF thermal analysis

### Files Created
1. **AUTOMATION_FRAMEWORK_PROMPTS.md** (850 lines, ~40 KB)
   - Complete automation framework with linearity focus
   - Structured prompts for Template/Guided/Autonomous phases
   - Algorithm pseudocode (sweet spot finder, GA optimizer)
   - Comprehensive report template

2. **linearity_optimizer.py** (850 lines, ~35 KB)
   - PA Simulator with simplified analytical models
   - Sweet Spot Finder (grid search: 50×50 = 2,500 points)
   - Multi-objective GA using differential evolution
   - Pareto Front Analyzer
   - 6-panel trade-off visualization tool
   - Outputs: `pa_tradeoff_analysis.png` (16×12 in, 300 DPI)

3. **plot_tradeoffs.R** (550 lines, ~20 KB)
   - Professional ggplot2 visualizations
   - Interactive 3D plotly (Pout × PAE × IM3)
   - Pareto front identification (R implementation)
   - 6 publication-quality plots + 1 interactive HTML
   - Outputs: `PA_Tradeoff_Analysis.png`, `PA_Tradeoff_3D.html`

4. **ENHANCEMENT_SUMMARY.md**
   - Complete documentation of enhancement
   - Usage instructions
   - Technical implementation details
   - Quality metrics

### Impact on Project
- **Linearity Priority**: Now a first-class consideration throughout design flow
- **Visual Decision Support**: Designers can see trade-offs before committing to design
- **Manufacturing-Aware Design**: Cost and yield considered from day one
- **Automation-Ready**: Structured prompts enable template/AI-based PA design
- **Reusability**: Scripts work as standalone tools or integrated modules

---

# 📝 Session Log: Week 2 Data Extraction & Chapter Development

## Session 2: February 1, 2026 - DATA EXTRACTION INITIATION ✅

### Completed Activities

#### 1. Data Extraction Framework Created ✅
**Files Created:**
- `Data_Extraction/Tx_Baseline_Extraction_Plan.md` (500+ lines)
  - Comprehensive analysis of 80 PDFs across 7 folders
  - Extraction methodology and templates defined
  - Priority High/Medium/Low assignments
  - Timeline: 15-20 hours over Weeks 2-7
  
**Key Findings:**
- **Background Overview** (4 PDFs): Project context and specifications
- **Design/Simulations** (40+ PDFs): 9 DOE builds systematically documented
  - DOE1 (12mm T9095A baseline)
  - DOE6 (Chip-n-wire 12mm T9095A)
  - DOE7 (2.4mm T9507B small device)
  - DOE9 (11.52mm R9505 large alternative)
  - DOE11 (3.84mm T9504 medium device)
  - DOE13 (6.4mm R9505 medium-large)
  - DOE15 (LDMOS T6083A technology comparison)
  - DOE16 (3.2mm R6051A optimized)
  - DOE17 (2.4mm T9501R latest technology)
- **Measurements** (15+ PDFs): Downselection reports and validation data
- **Modeling** (7+ PDFs): Device characterization
- **Assembly/Package** (8+ PDFs): Manufacturing procedures
- **Baseline Results** (1 PDF): Final selection documentation

#### 2. Extraction Templates Defined ✅
**Files Created:**
- `Data_Extraction/Tx_Baseline/01_Specifications.md` (400+ lines)
  - Template for extracting 5G frontend requirements
  - Performance targets: Frequency (3.3-3.8 GHz), Pout, PAE, IM3, ACPR, EVM
  - Design constraints and acceptance criteria
  - Integration roadmap with Chapter 1
  
- `Data_Extraction/Tx_Baseline/02_DOE_Build_Templates.md` (650+ lines)
  - Systematic extraction template for all 9 DOE builds
  - Device selection criteria, bias conditions, matching networks
  - Simulation results structure (S-parameters, load-pull, linearity)
  - Comparative analysis tables ready for population
  - Technology comparison framework (T-series vs R-series vs LDMOS)

**Extraction Methodology:**
- Manual PDF review with structured note-taking
- Per-PDF template capturing: Overview → Key Data → Figures → Insights → References
- Storage: Individual markdown files per DOE + consolidated comparison tables

#### 3. Chapter 1 RMarkdown Template Created ✅
**File Created:**
- `Chapters/Chapter_01_Transistor_Fundamentals.Rmd` (1,150+ lines, ~70 KB)
  - **Complete structure** with 8 major sections:
    1. Introduction (overview, learning outcomes, roadmap, prerequisites)
    2. GaN Device Physics (HEMT structure, 2DEG, DC characteristics)
    3. RF Characterization (S-parameters, load-pull, thermal)
    4. Transistor Downselection ⭐ CORE (project context, DOE methodology, selection criteria, results)
    5. Practical Design (biasing, matching, stability, layout)
    6. Modeling & Simulation (model types, extraction, validation)
    7. Reliability & Packaging (thermal, reliability screening)
    8. Summary & Conclusions (takeaways, checklist, transition to Chapter 2)
  
  - **Advanced Features**:
    - Tabbed sections (`.tabset`) for better navigation
    - R code chunks for data visualization (placeholders ready)
    - Kable tables for specifications and DOE comparison
    - Figures with captions and cross-references
    - Math equations in LaTeX/KaTeX
    - Bibliography integration
    - CSS styling for professional appearance
  
  - **Evidence Integration Points**:
    - 80 TBD placeholders for actual data from PDFs
    - Cross-references to IFX Tx_Baseline documents
    - Plot placeholders (load-pull contours, stability circles, DOE comparisons)
    - Measurement correlation tables
  
  - **Current Status**: 
    - ✅ Structure complete, ready for data population
    - 🔄 Data extraction in progress (0/80 PDFs processed)
    - ⏳ Figures to be generated from extracted data
    - ⏳ All "TBD" values to be filled from PDF review

#### 4. DEVELOPER_GUIDE.md Updated ✅
- Version bumped: 1.0 → 1.1
- Status updated: "Planning Phase" → "Active Development - Week 2"
- Reflects current data extraction focus

### Work in Progress

#### Data Extraction (In Progress) 🔄
**Status**: 0/80 PDFs extracted
**Current Focus**: 
- Next: Extract specifications from `5G_Frontend_requirements_PAM2p0+_external_2v1.pdf`
- Priority Week 2: Specifications + DOE1, DOE9, DOE11, DOE15, DOE17 (5 builds)
- Timeline: ~3-4 hours per DOE build = 15-20 hours for Week 2 target

**Folders Analyzed:**
1. `01_Background_overview/` (4 PDFs):
   - `01_Base_line_overview.pdf`
   - `02_Deep_dive_MUC_v0_general overview of Baseline.pdf`
   - `03_PAM2p0_Steering_2022-09-29_slides.pdf`
   - `04_TX_Baseline_overview_Jan_2023.pdf`

2. `02_Design/04_Specifications/` (1 PDF):
   - `5G_Frontend_requirements_PAM2p0+_external_2v1.pdf` ⭐ HIGH PRIORITY

3. `02_Design/06_Simulations/` (13 PDFs across 9 DOE builds):
   - Build structure and naming conventions documented
   - Size range: 2.4mm to 12mm (5× variation)
   - Technologies: T-series GaN, R-series GaN, LDMOS
   - Assembly variants: Standard wire-bond, Chip-and-wire

### Workspace Organization

**Project Structure Enhanced:**
```
PA_Design_Reference_Manual/
├── Data_Extraction/          ← NEW
│   ├── Tx_Baseline_Extraction_Plan.md
│   └── Tx_Baseline/
│       ├── 01_Specifications.md
│       ├── 02_DOE_Build_Templates.md
│       ├── 02_DOE_Builds/    ← Ready for individual DOE files
│       ├── 03_Measurements/
│       ├── 04_Modeling/
│       ├── 05_Assembly/
│       ├── 06_Package/
│       └── 07_Baseline_Results.md
├── Chapters/
│   ├── Chapter_01_Transistor_Fundamentals.Rmd  ← NEW (1150 lines)
│   ├── Chapter_01_Transistor_Fundamentals_OUTLINE.md (reference)
│   └── [Chapter 2-6 outlines exist]
├── DEVELOPER_GUIDE.md (v1.1)
├── PROJECT_STATUS.Rmd (this file, updated)
└── [Other project files...]
```

**Atomic_Cosmic_RFView_Project/**: 27 files organized in separate folder (completed earlier)

### Quality Metrics

- **Completeness**: 
  - ✅ 100% extraction methodology defined
  - ✅ 100% Chapter 1 structure complete
  - 🔄 0% data extracted (manual PDF review required)
  
- **Alignment with Plan**:
  - ✅ Week 2 deliverables on track
  - ✅ DEVELOPER_GUIDE complete
  - 🔄 Data extraction initiated (5/80 PDFs prioritized for Week 2)
  - ✅ Chapter 1 template ready
  - ⏳ Chapter 1 content awaits data extraction
  
- **Documentation Quality**:
  - All files have clear headers, purposes, and cross-references
  - Extraction templates systematic and repeatable
  - Chapter 1 follows RMarkdown best practices (tabbed sections, citations, code folding)

---

# 📝 Action Items for Next Check-in

## By February 5, 2026 (Week 2 Deadline)

### High Priority
- [ ] **Extract Specifications PDF** (est. 30-45 min)
  - Fill in all "TBD" values in `01_Specifications.md`
  - Capture key figures and page references
  
- [ ] **Extract 5 DOE Builds** (est. 15-20 hours total)
  - [ ] DOE1 (12mm T9095A baseline) - 3-4 hours
  - [ ] DOE9 (11.52mm R9505) - 3-4 hours
  - [ ] DOE11 (3.84mm T9504) - 3-4 hours
  - [ ] DOE15 (LDMOS baseline) - 3-4 hours
  - [ ] DOE17 (2.4mm T9501R latest) - 3-4 hours
  - Create individual `.md` files in `02_DOE_Builds/`
  
- [ ] **Populate Chapter 1 Section 1.4** (est. 2-3 hours)
  - Use extracted DOE data
  - Create comparison tables
  - Begin writing downselection narrative
  
- [ ] **Generate Initial Plots** (est. 1-2 hours)
  - DOE comparison: Size vs Pout, Size vs PAE
  - Technology comparison: GaN T-series vs R-series vs LDMOS
  - Use our `plot_tradeoffs.R` script with extracted data

### Medium Priority
- [ ] Update PROJECT_STATUS with extracted data summary
- [ ] Create extraction_progress.md tracking log
- [ ] Review and validate DEVELOPER_GUIDE completeness

### Stretch Goals (if time permits)
- [ ] Extract baseline results PDF (`07_PCB-pac/`)
- [ ] Begin DOE6, DOE7, DOE13, DOE16 extraction
- [ ] Start Chapter 1 Section 1.1 and 1.2 content writing

## Review Questions
1. Should extraction focus on performance metrics first, or include full design details?
2. Priority: Breadth (all 9 DOEs briefly) or Depth (5 DOEs thoroughly)? → **Depth chosen for Week 2**
3. Generate plots as we extract data, or wait until all DOEs complete?
4. Begin Chapter 1 writing now with partial data, or wait for more extraction?

---

# 📧 Contact & Updates

## How to Update This Document
1. Open `PROJECT_STATUS.Rmd` in RStudio/VS Code
2. Update relevant sections
3. Render: `rmarkdown::render("PROJECT_STATUS.Rmd")`
4. Commit changes to Git

## Quick Status Check
```{r quick_status, echo=TRUE, eval=FALSE}
# Run this in R console for quick status
cat(paste0(
  "Project: PA Design Reference Manual\n",
  "Phase: Setup & Planning\n",
  "Completion: 7.5%\n",
  "Next Milestone: Data Extraction (Feb 3-4)\n",
  "Status: ✅ On Track\n"
))
```

---

*Last Updated: `r Sys.Date()` | Auto-generated from PROJECT_STATUS.Rmd*
