---
title: "PA Class Integrated Visualization - Preview"
output:
  html_document:
    theme: cosmo
    toc: false
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Integrated PA Class Visualization

This combined visualization shows the I-V load line behavior alongside the corresponding input and output waveforms for different PA classes.

```{r pa-class-integrated-visualization, echo=FALSE, fig.cap="Figure 1.3d: Integrated PA Class Visualization - Load Lines and Input/Output Waveforms", fig.width=14, fig.height=10}
library(plotly)
library(dplyr)

# ========== DEVICE PARAMETERS ==========
VDD <- 48       # Supply voltage (V)
Vknee <- 4      # Knee voltage (V)
Idss <- 800     # Maximum current (mA)
Vth <- -5.5     # Threshold voltage (V)
Rload <- 55     # Load resistance (Ohms)
BVdss <- 100    # Breakdown voltage

# ========== GENERATE I-V CURVES ==========
VDS_curve <- seq(0.1, 100, 0.5)
VGS_values <- seq(-5.5, 0, 0.5)

iv_data <- data.frame()
for(VGS in VGS_values) {
  if(VGS < Vth) {
    ID <- rep(0, length(VDS_curve))
  } else {
    gm_factor <- (VGS - Vth) / (0 - Vth)
    ID_linear <- Idss * gm_factor * (VDS_curve / Vknee) * (1 - 0.5 * VDS_curve / Vknee)
    ID_sat <- Idss * gm_factor^2 * (1 + 0.01 * (VDS_curve - Vknee))
    ID <- ifelse(VDS_curve < Vknee, pmax(ID_linear, 0), ID_sat)
    ID <- ifelse(VDS_curve > BVdss, ID * (1 + 0.05*(VDS_curve - BVdss)), ID)
  }
  iv_data <- rbind(iv_data, data.frame(VDS = VDS_curve, ID = ID, VGS = VGS))
}

# ========== TIME DOMAIN WAVEFORMS ==========
t <- seq(0, 1, length.out = 200)
omega <- 2 * pi

# CLASS A: Full 360° conduction
VGS_Q_A <- -2.8
VGS_swing_A <- 0.8
VGS_A <- VGS_Q_A + VGS_swing_A * sin(omega * t)
gm_factor_A <- ((VGS_A - Vth) / (0 - Vth))^1.8
ID_A <- pmax(Idss * gm_factor_A, 0)
VDS_A <- pmax(pmin(VDD - ID_A * Rload, VDD), Vknee)

# CLASS AB: ~180-270° conduction
VGS_Q_AB <- -4.2
VGS_swing_AB <- 1.2
VGS_AB <- VGS_Q_AB + VGS_swing_AB * sin(omega * t)
gm_factor_AB <- ((VGS_AB - Vth) / (0 - Vth))^1.8
ID_AB <- pmax(Idss * gm_factor_AB, 0)
VDS_AB <- pmax(pmin(VDD - ID_AB * Rload, VDD), Vknee)

# CLASS B: 180° conduction
VGS_Q_B <- Vth
VGS_swing_B <- 1.5
VGS_B <- VGS_Q_B + VGS_swing_B * sin(omega * t)
gm_factor_B <- ((VGS_B - Vth) / (0 - Vth))^1.8
ID_B <- pmax(Idss * gm_factor_B, 0)
VDS_B <- pmax(pmin(VDD - ID_B * Rload, VDD), Vknee)

# CLASS C: <180° conduction
VGS_Q_C <- Vth - 1.0
VGS_swing_C <- 2.0
VGS_C <- VGS_Q_C + VGS_swing_C * sin(omega * t)
gm_factor_C <- ((VGS_C - Vth) / (0 - Vth))^1.8
ID_C <- pmax(Idss * gm_factor_C, 0)
VDS_C <- pmax(pmin(VDD - ID_C * Rload, VDD * 1.3), Vknee)

# CLASS D: Switching mode
VGS_Q_D <- Vth - 2.0
VGS_swing_D <- 3.0
VGS_D <- VGS_Q_D + VGS_swing_D * sin(omega * t)
ID_D <- ifelse(VGS_D > Vth, Idss * 0.7, 0)
VDS_D <- ifelse(ID_D > 0, Vknee, VDD)

# CLASS E: ZVS
VGS_Q_E <- Vth - 2.0
VGS_swing_E <- 3.5
VGS_E <- VGS_Q_E + VGS_swing_E * sin(omega * t)
ID_E <- ifelse(VGS_E > Vth, Idss * 0.6, 0)
phase_shift <- pi/6
VDS_E <- pmax(pmin(VDD * (1 + 0.8 * sin(omega * t + phase_shift)), VDD * 1.8), 0)

# CLASS F: Harmonic tuning
VGS_Q_F <- Vth - 1.5
VGS_swing_F <- 2.5
VGS_F <- VGS_Q_F + VGS_swing_F * sin(omega * t)
gm_factor_F <- ((VGS_F - Vth) / (0 - Vth))^1.8
ID_F <- pmax(Idss * gm_factor_F * 0.7, 0)
VDS_F <- pmax(pmin(VDD * (1 + 0.6 * sin(omega * t) + 0.2 * sin(3 * omega * t)), VDD * 1.5), Vknee)

# ========== LOAD LINE TRAJECTORIES ==========
# Class A Load Line
class_A_VDS <- c(Vknee, VDD/2, VDD)
class_A_ID <- c(Idss * 0.9, Idss * 0.45, 0)

# Class AB Load Line
class_AB_VDS <- c(Vknee, VDD/2, VDD)
class_AB_ID <- c(Idss * 0.85, Idss * 0.25, 0)

# Class B Load Line
class_B_VDS <- c(Vknee, VDD)
class_B_ID <- c(Idss * 0.7, 0)

# Class C Load Line
class_C_VDS <- c(Vknee, VDD * 1.1)
class_C_ID <- c(Idss * 0.5, 0)

# Class D Load Line
class_D_VDS <- c(Vknee, Vknee, VDD, VDD)
class_D_ID <- c(0, Idss * 0.8, Idss * 0.8, 0)

# Class E Load Line
class_E_VDS <- c(0.5, Vknee, VDD * 1.2, VDD * 1.8, 0.5)
class_E_ID <- c(Idss * 0.6, Idss * 0.7, Idss * 0.2, 0, Idss * 0.6)

# Class F Load Line
class_F_VDS <- seq(Vknee, VDD * 1.5, length.out = 30)
class_F_ID <- pmax(Idss * 0.4 * (1 + 0.6*sin(2*pi*seq(0,1,length.out=30))) * exp(-0.5*seq(0,1,length.out=30)), 0)

# ========== CREATE SUBPLOT 1: I-V CURVES WITH LOAD LINES ==========
p1 <- plot_ly()

# Add I-V curves (background, lighter)
for(VGS_val in VGS_values) {
  curve_data <- iv_data %>% filter(VGS == VGS_val)
  p1 <- p1 %>% add_trace(
    x = curve_data$VDS, y = curve_data$ID,
    type = 'scatter', mode = 'lines',
    line = list(width = 1, color = 'rgba(100,100,100,0.3)'),
    name = paste0('VGS=', VGS_val, 'V'),
    showlegend = FALSE,
    hovertemplate = paste0('VGS=', VGS_val, 'V<br>VDS: %{x:.1f}V<br>ID: %{y:.0f}mA<extra></extra>')
  )
}

# Add load lines for each class
p1 <- p1 %>%
  add_trace(x = class_A_VDS, y = class_A_ID, type = 'scatter', mode = 'lines+markers',
            name = 'Class A', legendgroup = 'classA',
            line = list(color = '#FF6F00', width = 3, dash = 'dash'),
            marker = list(size = 8, symbol = 'circle'),
            hovertemplate = '<b>Class A</b><br>VDS: %{x:.1f}V<br>ID: %{y:.0f}mA<extra></extra>') %>%
  add_trace(x = class_AB_VDS, y = class_AB_ID, type = 'scatter', mode = 'lines+markers',
            name = 'Class AB', legendgroup = 'classAB', visible = 'legendonly',
            line = list(color = '#4CAF50', width = 3, dash = 'dashdot'),
            marker = list(size = 8, symbol = 'square'),
            hovertemplate = '<b>Class AB</b><br>VDS: %{x:.1f}V<br>ID: %{y:.0f}mA<extra></extra>') %>%
  add_trace(x = class_B_VDS, y = class_B_ID, type = 'scatter', mode = 'lines+markers',
            name = 'Class B', legendgroup = 'classB', visible = 'legendonly',
            line = list(color = '#2196F3', width = 3, dash = 'dot'),
            marker = list(size = 8, symbol = 'diamond'),
            hovertemplate = '<b>Class B</b><br>VDS: %{x:.1f}V<br>ID: %{y:.0f}mA<extra></extra>') %>%
  add_trace(x = class_C_VDS, y = class_C_ID, type = 'scatter', mode = 'lines+markers',
            name = 'Class C', legendgroup = 'classC', visible = 'legendonly',
            line = list(color = '#9C27B0', width = 3, dash = 'longdash'),
            marker = list(size = 8, symbol = 'triangle-up'),
            hovertemplate = '<b>Class C</b><br>VDS: %{x:.1f}V<br>ID: %{y:.0f}mA<extra></extra>') %>%
  add_trace(x = class_D_VDS, y = class_D_ID, type = 'scatter', mode = 'lines+markers',
            name = 'Class D', legendgroup = 'classD', visible = 'legendonly',
            line = list(color = '#FF5722', width = 3, dash = 'solid'),
            marker = list(size = 8, symbol = 'x'),
            hovertemplate = '<b>Class D</b><br>VDS: %{x:.1f}V<br>ID: %{y:.0f}mA<extra></extra>') %>%
  add_trace(x = class_E_VDS, y = class_E_ID, type = 'scatter', mode = 'lines+markers',
            name = 'Class E', legendgroup = 'classE', visible = 'legendonly',
            line = list(color = '#E91E63', width = 3, dash = 'dashdot'),
            marker = list(size = 8, symbol = 'star'),
            hovertemplate = '<b>Class E</b><br>VDS: %{x:.1f}V<br>ID: %{y:.0f}mA<extra></extra>') %>%
  add_trace(x = class_F_VDS, y = class_F_ID, type = 'scatter', mode = 'lines+markers',
            name = 'Class F', legendgroup = 'classF', visible = 'legendonly',
            line = list(color = '#00BCD4', width = 3, dash = 'dot'),
            marker = list(size = 7, symbol = 'hexagon'),
            hovertemplate = '<b>Class F</b><br>VDS: %{x:.1f}V<br>ID: %{y:.0f}mA<extra></extra>')

# Add reference lines
p1 <- p1 %>%
  add_segments(x = Vknee, xend = Vknee, y = 0, yend = 900,
               line = list(color = 'gray', width = 1, dash = 'dash'),
               showlegend = FALSE, hovertemplate = paste0('Vknee = ', Vknee, 'V<extra></extra>')) %>%
  add_segments(x = VDD, xend = VDD, y = 0, yend = 900,
               line = list(color = 'red', width = 1, dash = 'dash'),
               showlegend = FALSE, hovertemplate = paste0('VDD = ', VDD, 'V<extra></extra>'))

p1 <- p1 %>% layout(
  xaxis = list(title = '<b>V<sub>DS</sub> [V]</b>', range = c(0, 90)),
  yaxis = list(title = '<b>I<sub>D</sub> [mA]</b>', range = c(0, 900)),
  annotations = list(
    list(x = 0.5, y = 1.05, xref = 'paper', yref = 'paper',
         text = '<b>Load Lines on I-V Curves</b>',
         showarrow = FALSE, font = list(size = 13))
  )
)

# ========== SUBPLOT 2: INPUT VGS WAVEFORMS ==========
p2 <- plot_ly() %>%
  add_trace(x = t, y = VGS_A, name = 'Class A', type = 'scatter', mode = 'lines',
            legendgroup = 'classA', showlegend = FALSE,
            line = list(color = '#FF6F00', width = 2.5),
            hovertemplate = '<b>Class A</b><br>t: %{x:.3f}<br>VGS: %{y:.2f}V<extra></extra>') %>%
  add_trace(x = t, y = VGS_AB, name = 'Class AB', type = 'scatter', mode = 'lines',
            legendgroup = 'classAB', visible = 'legendonly', showlegend = FALSE,
            line = list(color = '#4CAF50', width = 2.5),
            hovertemplate = '<b>Class AB</b><br>t: %{x:.3f}<br>VGS: %{y:.2f}V<extra></extra>') %>%
  add_trace(x = t, y = VGS_B, name = 'Class B', type = 'scatter', mode = 'lines',
            legendgroup = 'classB', visible = 'legendonly', showlegend = FALSE,
            line = list(color = '#2196F3', width = 2.5),
            hovertemplate = '<b>Class B</b><br>t: %{x:.3f}<br>VGS: %{y:.2f}V<extra></extra>') %>%
  add_trace(x = t, y = VGS_C, name = 'Class C', type = 'scatter', mode = 'lines',
            legendgroup = 'classC', visible = 'legendonly', showlegend = FALSE,
            line = list(color = '#9C27B0', width = 2.5),
            hovertemplate = '<b>Class C</b><br>t: %{x:.3f}<br>VGS: %{y:.2f}V<extra></extra>') %>%
  add_trace(x = t, y = VGS_D, name = 'Class D', type = 'scatter', mode = 'lines',
            legendgroup = 'classD', visible = 'legendonly', showlegend = FALSE,
            line = list(color = '#FF5722', width = 2.5),
            hovertemplate = '<b>Class D</b><br>t: %{x:.3f}<br>VGS: %{y:.2f}V<extra></extra>') %>%
  add_trace(x = t, y = VGS_E, name = 'Class E', type = 'scatter', mode = 'lines',
            legendgroup = 'classE', visible = 'legendonly', showlegend = FALSE,
            line = list(color = '#E91E63', width = 2.5),
            hovertemplate = '<b>Class E</b><br>t: %{x:.3f}<br>VGS: %{y:.2f}V<extra></extra>') %>%
  add_trace(x = t, y = VGS_F, name = 'Class F', type = 'scatter', mode = 'lines',
            legendgroup = 'classF', visible = 'legendonly', showlegend = FALSE,
            line = list(color = '#00BCD4', width = 2.5),
            hovertemplate = '<b>Class F</b><br>t: %{x:.3f}<br>VGS: %{y:.2f}V<extra></extra>') %>%
  add_segments(x = 0, xend = 1, y = Vth, yend = Vth,
               line = list(color = 'gray', width = 1, dash = 'dash'),
               showlegend = FALSE, hovertemplate = paste0('Vth = ', Vth, 'V<extra></extra>'))

p2 <- p2 %>% layout(
  xaxis = list(title = '', showticklabels = FALSE),
  yaxis = list(title = '<b>V<sub>GS</sub> Input [V]</b>'),
  annotations = list(
    list(x = 0.5, y = 1.05, xref = 'paper', yref = 'paper',
         text = '<b>Gate Input Waveform</b>',
         showarrow = FALSE, font = list(size = 13))
  )
)

# ========== SUBPLOT 3: OUTPUT VDS WAVEFORMS ==========
p3 <- plot_ly() %>%
  add_trace(x = t, y = VDS_A, name = 'Class A', type = 'scatter', mode = 'lines',
            legendgroup = 'classA', showlegend = FALSE,
            line = list(color = '#FF6F00', width = 2.5),
            hovertemplate = '<b>Class A</b><br>t: %{x:.3f}<br>VDS: %{y:.1f}V<extra></extra>') %>%
  add_trace(x = t, y = VDS_AB, name = 'Class AB', type = 'scatter', mode = 'lines',
            legendgroup = 'classAB', visible = 'legendonly', showlegend = FALSE,
            line = list(color = '#4CAF50', width = 2.5),
            hovertemplate = '<b>Class AB</b><br>t: %{x:.3f}<br>VDS: %{y:.1f}V<extra></extra>') %>%
  add_trace(x = t, y = VDS_B, name = 'Class B', type = 'scatter', mode = 'lines',
            legendgroup = 'classB', visible = 'legendonly', showlegend = FALSE,
            line = list(color = '#2196F3', width = 2.5),
            hovertemplate = '<b>Class B</b><br>t: %{x:.3f}<br>VDS: %{y:.1f}V<extra></extra>') %>%
  add_trace(x = t, y = VDS_C, name = 'Class C', type = 'scatter', mode = 'lines',
            legendgroup = 'classC', visible = 'legendonly', showlegend = FALSE,
            line = list(color = '#9C27B0', width = 2.5),
            hovertemplate = '<b>Class C</b><br>t: %{x:.3f}<br>VDS: %{y:.1f}V<extra></extra>') %>%
  add_trace(x = t, y = VDS_D, name = 'Class D', type = 'scatter', mode = 'lines',
            legendgroup = 'classD', visible = 'legendonly', showlegend = FALSE,
            line = list(color = '#FF5722', width = 2.5),
            hovertemplate = '<b>Class D</b><br>t: %{x:.3f}<br>VDS: %{y:.1f}V<extra></extra>') %>%
  add_trace(x = t, y = VDS_E, name = 'Class E', type = 'scatter', mode = 'lines',
            legendgroup = 'classE', visible = 'legendonly', showlegend = FALSE,
            line = list(color = '#E91E63', width = 2.5),
            hovertemplate = '<b>Class E</b><br>t: %{x:.3f}<br>VDS: %{y:.1f}V<extra></extra>') %>%
  add_trace(x = t, y = VDS_F, name = 'Class F', type = 'scatter', mode = 'lines',
            legendgroup = 'classF', visible = 'legendonly', showlegend = FALSE,
            line = list(color = '#00BCD4', width = 2.5),
            hovertemplate = '<b>Class F</b><br>t: %{x:.3f}<br>VDS: %{y:.1f}V<extra></extra>') %>%
  add_segments(x = 0, xend = 1, y = VDD, yend = VDD,
               line = list(color = 'red', width = 1, dash = 'dash'),
               showlegend = FALSE, hovertemplate = paste0('VDD = ', VDD, 'V<extra></extra>')) %>%
  add_segments(x = 0, xend = 1, y = Vknee, yend = Vknee,
               line = list(color = 'gray', width = 1, dash = 'dash'),
               showlegend = FALSE, hovertemplate = paste0('Vknee = ', Vknee, 'V<extra></extra>'))

p3 <- p3 %>% layout(
  xaxis = list(title = '<b>Normalized Time (RF Cycle)</b>'),
  yaxis = list(title = '<b>V<sub>DS</sub> Output [V]</b>'),
  annotations = list(
    list(x = 0.5, y = 1.05, xref = 'paper', yref = 'paper',
         text = '<b>Drain Voltage Waveform</b>',
         showarrow = FALSE, font = list(size = 13))
  )
)

# ========== SUBPLOT 4: OUTPUT ID WAVEFORMS ==========
p4 <- plot_ly() %>%
  add_trace(x = t, y = ID_A, name = 'Class A', type = 'scatter', mode = 'lines',
            legendgroup = 'classA', showlegend = FALSE,
            line = list(color = '#FF6F00', width = 2.5),
            hovertemplate = '<b>Class A</b><br>t: %{x:.3f}<br>ID: %{y:.0f}mA<extra></extra>') %>%
  add_trace(x = t, y = ID_AB, name = 'Class AB', type = 'scatter', mode = 'lines',
            legendgroup = 'classAB', visible = 'legendonly', showlegend = FALSE,
            line = list(color = '#4CAF50', width = 2.5),
            hovertemplate = '<b>Class AB</b><br>t: %{x:.3f}<br>ID: %{y:.0f}mA<extra></extra>') %>%
  add_trace(x = t, y = ID_B, name = 'Class B', type = 'scatter', mode = 'lines',
            legendgroup = 'classB', visible = 'legendonly', showlegend = FALSE,
            line = list(color = '#2196F3', width = 2.5),
            hovertemplate = '<b>Class B</b><br>t: %{x:.3f}<br>ID: %{y:.0f}mA<extra></extra>') %>%
  add_trace(x = t, y = ID_C, name = 'Class C', type = 'scatter', mode = 'lines',
            legendgroup = 'classC', visible = 'legendonly', showlegend = FALSE,
            line = list(color = '#9C27B0', width = 2.5),
            hovertemplate = '<b>Class C</b><br>t: %{x:.3f}<br>ID: %{y:.0f}mA<extra></extra>') %>%
  add_trace(x = t, y = ID_D, name = 'Class D', type = 'scatter', mode = 'lines',
            legendgroup = 'classD', visible = 'legendonly', showlegend = FALSE,
            line = list(color = '#FF5722', width = 2.5),
            hovertemplate = '<b>Class D</b><br>t: %{x:.3f}<br>ID: %{y:.0f}mA<extra></extra>') %>%
  add_trace(x = t, y = ID_E, name = 'Class E', type = 'scatter', mode = 'lines',
            legendgroup = 'classE', visible = 'legendonly', showlegend = FALSE,
            line = list(color = '#E91E63', width = 2.5),
            hovertemplate = '<b>Class E</b><br>t: %{x:.3f}<br>ID: %{y:.0f}mA<extra></extra>') %>%
  add_trace(x = t, y = ID_F, name = 'Class F', type = 'scatter', mode = 'lines',
            legendgroup = 'classF', visible = 'legendonly', showlegend = FALSE,
            line = list(color = '#00BCD4', width = 2.5),
            hovertemplate = '<b>Class F</b><br>t: %{x:.3f}<br>ID: %{y:.0f}mA<extra></extra>')

p4 <- p4 %>% layout(
  xaxis = list(title = '<b>Normalized Time (RF Cycle)</b>'),
  yaxis = list(title = '<b>I<sub>D</sub> Output [mA]</b>'),
  annotations = list(
    list(x = 0.5, y = 1.05, xref = 'paper', yref = 'paper',
         text = '<b>Drain Current Waveform</b>',
         showarrow = FALSE, font = list(size = 13))
  )
)

# ========== COMBINE INTO 2x2 GRID ==========
fig <- subplot(
  p1, p2, p3, p4,
  nrows = 2,
  margin = 0.08,
  shareX = FALSE,
  shareY = FALSE
) %>% layout(
  title = list(
    text = '<b>Integrated PA Class Visualization: Load Lines ↔ Input/Output Waveforms</b><br><sub>Click legend to toggle PA classes • Class A shown by default</sub>',
    font = list(size = 16),
    y = 0.98
  ),
  width = 1400,
  height = 1000,
  hovermode = 'closest',
  plot_bgcolor = 'white',
  margin = list(l = 60, r = 120, t = 100, b = 60),
  legend = list(
    orientation = 'v',
    x = 1.01,
    y = 1,
    bgcolor = 'rgba(255,255,255,0.95)',
    bordercolor = 'gray',
    borderwidth = 2,
    font = list(size = 11)
  ),
  showlegend = TRUE
)

fig
```

---

## How to Use This Integrated Visualization

### Layout (2×2 Grid):
- **Top Left**: I-V curves with load line trajectories
- **Top Right**: Input gate voltage (VGS) waveforms  
- **Bottom Left**: Output drain voltage (VDS) waveforms
- **Bottom Right**: Output drain current (ID) waveforms

### Interactive Controls:
1. Click any **PA class** in the legend to show/hide all four related plots
2. **Class A** is shown by default
3. Display **multiple classes** simultaneously for comparison
4. **Hover** over any waveform or load line to see exact values

### Understanding the Relationships:
- **Input (VGS)** drives the gate → modulates channel conductance
- **Drain current (ID)** responds to gate voltage based on transfer characteristic
- **Load line** shows the constraint: `VDS = VDD - ID × RL`
- **Operating point** moves along load line as ID varies
- **VDS and ID are anti-phase**: When ID ↑, VDS ↓ (visible in bottom plots)

### Key Insights by PA Class:
- **Class A**: 360° conduction, symmetric waveforms, centered Q-point
- **Class AB**: Partial clipping visible in ID waveform, 180-270° conduction
- **Class B**: Half-wave rectified ID (180° conduction), biased at threshold
- **Class C**: Narrow ID pulses (<180° conduction), voltage overshoot possible
- **Class D/E**: Hard switching, square waveforms
- **Class F**: Harmonic-shaped waveforms for efficiency

---

**This integrated view makes the load line behavior much clearer!** You can see how:
1. Input signal swing varies by class
2. Current waveform shape determines conduction angle
3. Operating point trajectory follows the load line
4. Output voltage is inverted and scaled version of drain current
