---
title: "Document Title"
subtitle: "Subtitle or Project Name"
author: "Author Name"
date: "`r format(Sys.Date(), '%d %B %Y')`"

# ── Output formats ────────────────────────────────────────────────────────────
# Render both: rmarkdown::render("document.Rmd", output_format = "all")
output:
  html_document:
    theme: flatly           # base Bootstrap theme
    highlight: kate
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    number_sections: true
    code_folding: hide      # show/hide toggle per chunk
    fig_caption: true
    fig_width: 7
    fig_height: 4.5
    df_print: kable
    includes:
      in_header: "../assets/rmd-header.html"    # fonts + day/night JS
    css: "../assets/rmd-custom.css"
  pdf_document:
    toc: true
    toc_depth: 3
    number_sections: true
    fig_caption: true
    fig_width: 6.5
    fig_height: 4
    latex_engine: xelatex
    includes:
      in_header: "../assets/rmd-latex-header.tex"
    citation_package: natbib

# ── Bibliography ──────────────────────────────────────────────────────────────
bibliography: references.bib
csl: https://www.zotero.org/styles/apa-7th-edition
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo      = TRUE,
  warning   = FALSE,
  message   = FALSE,
  fig.align = "center"
)

# Color-blind-safe Okabe-Ito palette
okabe_ito <- c(
  "#E69F00", "#56B4E9", "#009E73", "#F0E442",
  "#0072B2", "#D55E00", "#CC79A7", "#000000"
)

if (requireNamespace("ggplot2", quietly = TRUE)) {
  options(
    ggplot2.discrete.colour = okabe_ito,
    ggplot2.discrete.fill   = okabe_ito
  )
}

options(knitr.kable.NA = "—")
```

## Abstract {-}

Replace this block with a 150–250 word abstract.

---

## Introduction

Write your introduction here. Cite with `[@key2024]`.

### Background

Subsection content.

### Objectives

1. First objective
2. Second objective

---

## Methods {.tabset .tabset-fade .tabset-pills}

### Approach

Describe the overall approach.

### Data Sources

```{r data-summary}
df <- data.frame(
  Source  = c("Source A", "Source B"),
  Records = c(1200, 850),
  Period  = c("2020–2024", "2021–2024")
)
knitr::kable(df, align = "lrr", caption = "Summary of data sources.")
```

### Statistical Analysis

Describe the analysis.

---

## Results {.tabset .tabset-fade .tabset-pills}

### Overview

Summary of key findings.

### Charts

```{r fig-example, fig.cap="Example chart using the Okabe-Ito palette."}
if (requireNamespace("ggplot2", quietly = TRUE)) {
  library(ggplot2)
  example_data <- data.frame(
    category = c("Group A", "Group B", "Group C"),
    value    = c(42, 67, 31)
  )
  ggplot(example_data, aes(x = category, y = value, fill = category)) +
    geom_col(width = 0.6, show.legend = FALSE) +
    scale_fill_manual(values = okabe_ito[1:3]) +
    labs(x = NULL, y = "Value") +
    theme_minimal(base_size = 13) +
    theme(panel.grid.major.x = element_blank())
}
```

### Tables

```{r tbl-results}
results <- data.frame(
  Metric = c("Accuracy", "Precision", "Recall"),
  Value  = c(0.94, 0.91, 0.96),
  CI_95  = c("[0.91, 0.97]", "[0.88, 0.94]", "[0.93, 0.99]")
)
knitr::kable(results, digits = 2,
             col.names = c("Metric", "Value", "95% CI"),
             caption = "Key results summary.")
```

---

## Discussion

Discuss findings here.

### Limitations

> **Limitations**: Note any important caveats here.

---

## Conclusion

Summarise in 3–5 sentences.

---

## References {-}

<div id="refs"></div>

---

## Appendix {-}

### Session Info {-}

```{r session-info}
sessionInfo()
```
