---
title: "Document Title"
subtitle: "Subtitle or Project Name"
author:
  - name: "Author Name"
    affiliation: "Organisation"
    email: "author@example.com"
date: today
date-format: "D MMMM YYYY"

# ── Output formats ──────────────────────────────────────────────────────────
format:
  html:
    theme:
      light: [flatly,  ../assets/theme-light.scss]
      dark:  [darkly,  ../assets/theme-dark.scss]
    toc: true
    toc-depth: 3
    toc-location: left
    toc-title: "Contents"
    toc-expand: 2
    number-sections: true
    smooth-scroll: true
    code-fold: true
    code-tools: true
    code-copy: true
    fig-align: center
    fig-cap-location: bottom
    fig-responsive: true
    link-external-newwindow: true
    link-external-icon: true
    html-math-method: mathjax
    page-layout: article
    grid:
      sidebar-width: 260px
      body-width: 860px
    include-in-header:
      - text: |
          <meta name="color-scheme" content="light dark">
          <link rel="preconnect" href="https://fonts.googleapis.com">
          <link rel="preconnect" href="https://fonts.gstatic.com" crossorigin>
          <link href="https://fonts.googleapis.com/css2?family=Inter:wght@400;500;600;700&family=JetBrains+Mono:wght@400;500&display=swap" rel="stylesheet">
  pdf:
    documentclass: article
    geometry: "a4paper, margin=2.5cm"
    toc: true
    toc-depth: 3
    number-sections: true
    colorlinks: true
    linkcolor: "0072B2"
    urlcolor:  "0072B2"
    citecolor: "D55E00"
    fig-pos: "H"
    fig-align: center
    fig-cap-location: bottom
    keep-tex: false
    latex-engine: xelatex
    include-in-header:
      - text: |
          \usepackage{booktabs}
          \usepackage{longtable}
          \usepackage{float}
          \usepackage{hyperref}

# ── Bibliography ─────────────────────────────────────────────────────────────
bibliography: references.bib
csl: https://www.zotero.org/styles/apa-7th-edition
link-citations: true

# ── Global chunk options ──────────────────────────────────────────────────────
execute:
  echo: true
  warning: false
  message: false
  cache: false

# ── Cross-references ──────────────────────────────────────────────────────────
crossref:
  fig-title: "Figure"
  tbl-title: "Table"
  eq-prefix: "Equation"
---

```{r}
#| label: setup
#| include: false

# Color-blind-safe Okabe-Ito palette (8 colors)
okabe_ito <- c(
  orange        = "#E69F00",
  sky_blue      = "#56B4E9",
  bluish_green  = "#009E73",
  yellow        = "#F0E442",
  blue          = "#0072B2",
  vermillion    = "#D55E00",
  reddish_purple= "#CC79A7",
  black         = "#000000"
)

# Apply to all ggplot2 charts globally
if (requireNamespace("ggplot2", quietly = TRUE)) {
  options(
    ggplot2.discrete.colour = unname(okabe_ito),
    ggplot2.discrete.fill   = unname(okabe_ito)
  )
}

# Knitr table formatting
if (requireNamespace("knitr", quietly = TRUE)) {
  options(knitr.kable.NA = "—")
}
```

## Abstract {.unnumbered}

::: {.callout-note appearance="minimal"}
Replace this block with a 150–250 word abstract summarising the purpose,
methods, key findings, and conclusions of this document.
:::

---

## Introduction

Write your introduction here. Use `[@key2024]` to cite references.

### Background

Subsection content.

### Objectives

1. First objective
2. Second objective
3. Third objective

---

## Methods {.tabset}

### Approach

Describe the overall approach here.

### Data Sources

Describe data sources here.

```{r}
#| label: tbl-data-summary
#| tbl-cap: "Summary of data sources"

# Example table — replace with actual data
df <- data.frame(
  Source    = c("Source A", "Source B"),
  Records   = c(1200,       850),
  Period    = c("2020–2024", "2021–2024")
)

knitr::kable(df, align = "lrr", booktabs = TRUE)
```

### Statistical Analysis

Describe the analysis approach.

---

## Results {.tabset}

### Overview

Summary of key findings.

### Charts

```{r}
#| label: fig-example
#| fig-cap: "Example chart using the Okabe-Ito color-blind-safe palette."
#| fig-alt: "Bar chart with three categories using orange, sky-blue, and bluish-green bars."
#| fig-width: 7
#| fig-height: 4

if (requireNamespace("ggplot2", quietly = TRUE)) {
  library(ggplot2)

  example_data <- data.frame(
    category = c("Group A", "Group B", "Group C"),
    value    = c(42, 67, 31)
  )

  ggplot(example_data, aes(x = category, y = value, fill = category)) +
    geom_col(width = 0.6, show.legend = FALSE) +
    scale_fill_manual(values = unname(okabe_ito[1:3])) +
    labs(x = NULL, y = "Value") +
    theme_minimal(base_size = 13) +
    theme(panel.grid.major.x = element_blank())
}
```

### Tables

```{r}
#| label: tbl-results
#| tbl-cap: "Key results summary."

results <- data.frame(
  Metric    = c("Accuracy", "Precision", "Recall"),
  Value     = c(0.94,       0.91,        0.96),
  CI_95     = c("[0.91, 0.97]", "[0.88, 0.94]", "[0.93, 0.99]")
)

knitr::kable(results, digits = 2, col.names = c("Metric", "Value", "95% CI"),
             align = "lrr", booktabs = TRUE)
```

---

## Discussion

Discuss your findings. Cross-reference figures with @fig-example and tables with @tbl-results.

### Limitations

::: {.callout-warning}
**Limitations**: Note any important limitations of this work here.
:::

### Future Work

Describe next steps.

---

## Conclusion

Summarise key takeaways in 3–5 sentences.

---

## References {.unnumbered}

::: {#refs}
:::

---

## Appendix {.unnumbered .appendix}

### Appendix A: Additional Tables {.unnumbered}

Supplementary material here.

### Appendix B: Session Info {.unnumbered}

```{r}
#| label: session-info
#| code-fold: true
#| code-summary: "Show session info"

sessionInfo()
```
