---
title: "Chapter XX: [Chapter Title]"
author: "PA Design Reference Manual"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    number_sections: true
    theme: cosmo
    highlight: tango
    code_folding: hide
    fig_caption: true
    self_contained: true
    css: ../../shared_resources/styles/main.css
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width = 8,
  fig.height = 6,
  fig.align = "center",
  cache = FALSE
)

# Load required libraries
library(knitr)
library(kableExtra)
library(ggplot2)
library(dplyr)
library(here)

# Source utility functions
source(here("shared_resources/scripts/utils.R"))
source(here("shared_resources/scripts/plotting.R"))
```

---

# Introduction {.tabset .tabset-fade .tabset-pills}

## Motivation

**Why This Chapter Matters**

[Explain the practical importance of this topic in PA design]

**Real-World Context**

[Connect to actual project experience from IFX]

**Learning Objectives**

By the end of this chapter, you will understand:

- [Objective 1]
- [Objective 2]
- [Objective 3]

## Overview

**Chapter Structure**

This chapter follows a systematic progression:

1. **Theoretical Foundation** - Core concepts and equations
2. **Design Methodology** - Practical design approach
3. **Evidence from Projects** - Real data and measurements
4. **Current Trends** - State of the art
5. **Future Directions** - Emerging techniques
6. **Lessons Learned** - Practical insights

## Key Concepts

```{r concept-table}
concepts <- data.frame(
  Concept = c(
    "Concept 1",
    "Concept 2",
    "Concept 3"
  ),
  Description = c(
    "Brief description",
    "Brief description",
    "Brief description"
  ),
  Importance = c(
    "High",
    "Medium",
    "High"
  )
)

kable(concepts, caption = "Key Concepts Covered") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

# Theoretical Foundation {.tabset .tabset-fade .tabset-pills}

## Fundamental Principles

### Core Theory

[Present the theoretical foundation with equations]

**Key Equation:**

$$
[LaTeX equation]
$$

where:
- $variable$ = description
- $variable$ = description

### Physical Interpretation

[Explain the physics behind the theory]

```{r theory-plot, fig.cap="Theoretical Relationship"}
# Example plot
x <- seq(0, 10, 0.1)
y <- sin(x)

plot(x, y, type = "l", 
     xlab = "Parameter", 
     ylab = "Response",
     main = "Theoretical Behavior")
```

## Mathematical Framework

### Derivation

[Step-by-step derivation if relevant]

### Assumptions and Limitations

- **Assumption 1**: [Description and validity range]
- **Assumption 2**: [Description and validity range]
- **Limitation 1**: [When theory breaks down]

## Literature Review

### Seminal Work

Key references in this area:

- [@reference1] - First major contribution
- [@reference2] - Important development
- [@reference3] - Recent advances

### Current State of Art

[Summary of latest research and industry practices]

# Design Methodology {.tabset .tabset-fade .tabset-pills}

## Step-by-Step Process

### Design Flow

```{r design-flow}
# Create design flow diagram or table
flow <- data.frame(
  Step = 1:5,
  Phase = c(
    "Requirements Definition",
    "Initial Design",
    "Simulation",
    "Optimization",
    "Verification"
  ),
  Key_Activities = c(
    "Define specs, constraints",
    "Topology selection, preliminary sizing",
    "EM/circuit simulation",
    "Parameter tuning, trade-offs",
    "Performance validation"
  ),
  Deliverable = c(
    "Spec document",
    "Initial schematic",
    "Simulation results",
    "Optimized design",
    "Verified design"
  )
)

kable(flow, caption = "Design Flow") %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  column_spec(2, bold = TRUE)
```

### Practical Considerations

**Design Trade-offs:**

- Performance vs. Complexity
- Cost vs. Performance
- Time-to-market vs. Optimization

## Design Tools

### Simulation Tools

- **Tool 1**: [Usage and strengths]
- **Tool 2**: [Usage and strengths]

### Measurement Equipment

- **Equipment 1**: [Purpose and specifications]
- **Equipment 2**: [Purpose and specifications]

## Design Examples

### Example 1: [Specific Case]

**Requirements:**
- [Requirement 1]
- [Requirement 2]

**Solution Approach:**
[Step-by-step solution]

```{r example-1}
# Example calculations or plots
# Using data from IFX projects
```

# Evidence from IFX Projects {.tabset .tabset-fade .tabset-pills}

## Project Overview

**Source Project**: [IFX Project Name]  
**Timeframe**: [Date range]  
**Objective**: [Project goal]  
**Outcome**: [Result achieved]

### Project Context

[Background information from IFX project]

## Simulation Results

### Setup

**Simulation Tool**: [ADS/SystemVue/etc.]  
**Configuration**: [Key parameters]  
**Goal**: [What was being investigated]

### Results

```{r simulation-results}
# Load and plot simulation data from IFX project
# data <- read_csv(here("manual_chapters/chXX_[topic]/data/sim_results.csv"))

# plot_smith_chart(data, title = "Simulated Load-Pull Contours")
```

**Key Findings:**

1. [Finding 1 with supporting data]
2. [Finding 2 with supporting data]
3. [Finding 3 with supporting data]

## Measurement Data

### Test Setup

**Measurement Equipment**:
- [Equipment list]

**Test Conditions**:
- [Conditions]

### Measured Performance

```{r measurements}
# Load and plot measurement data
# measurement_data <- read_csv(here("manual_chapters/chXX_[topic]/data/measurements.csv"))

# Create plots comparing simulation vs measurement
```

**Comparison: Simulation vs. Measurement**

```{r sim-vs-meas-table}
comparison <- data.frame(
  Parameter = c("Parameter 1", "Parameter 2", "Parameter 3"),
  Simulated = c("Value", "Value", "Value"),
  Measured = c("Value", "Value", "Value"),
  Agreement = c("Excellent", "Good", "Excellent")
)

kable(comparison, caption = "Simulation vs. Measurement") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## Design Decisions & Rationale

### Key Decisions Made

**Decision 1: [Topic]**

- **Rationale**: [Why this choice was made]
- **Trade-offs**: [What was sacrificed]
- **Outcome**: [Result of the decision]
- **Lesson**: [What was learned]

**Decision 2: [Topic]**

- **Rationale**: [Why this choice was made]
- **Trade-offs**: [What was sacrificed]
- **Outcome**: [Result of the decision]
- **Lesson**: [What was learned]

# Trends & Current Status {.tabset .tabset-fade .tabset-pills}

## Industry Trends

### Current Trends

1. **Trend 1**: [Description and implications]
2. **Trend 2**: [Description and implications]
3. **Trend 3**: [Description and implications]

### Emerging Technologies

- [Technology 1]: [Maturity level and potential]
- [Technology 2]: [Maturity level and potential]

## State of the Art

### Performance Benchmarks

```{r sota-benchmarks}
# Comparison table of current state-of-art
sota <- data.frame(
  Metric = c("Metric 1", "Metric 2", "Metric 3"),
  State_of_Art = c("Value", "Value", "Value"),
  IFX_Achievement = c("Value", "Value", "Value"),
  Gap = c("+/-X%", "+/-X%", "+/-X%")
)

kable(sota, caption = "State of the Art Comparison") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

### Technology Readiness

[Assessment of maturity for different approaches]

# Future Directions {.tabset .tabset-fade .tabset-pills}

## Research Opportunities

### Open Problems

1. **Problem 1**: [Description and significance]
2. **Problem 2**: [Description and significance]

### Potential Solutions

- **Approach 1**: [Feasibility and challenges]
- **Approach 2**: [Feasibility and challenges]

## Technology Roadmap

### Near-term (1-2 years)

- [Expected development 1]
- [Expected development 2]

### Mid-term (3-5 years)

- [Anticipated advancement 1]
- [Anticipated advancement 2]

### Long-term (5+ years)

- [Future vision 1]
- [Future vision 2]

## Recommendations

**For New Designs:**

- [Recommendation 1]
- [Recommendation 2]

**For Research:**

- [Research direction 1]
- [Research direction 2]

# Lessons Learned {.tabset .tabset-fade .tabset-pills}

## Technical Lessons

### What Worked Well

1. **[Topic]**: [Successful approach and why]
2. **[Topic]**: [Successful approach and why]

### What Didn't Work

1. **[Topic]**: [Failed approach and why]
2. **[Topic]**: [Failed approach and why]

### Insights for Future

- [Actionable insight 1]
- [Actionable insight 2]

## Process Lessons

### Design Process

- [Process improvement 1]
- [Process improvement 2]

### Team Collaboration

- [Collaboration lesson 1]
- [Collaboration lesson 2]

## Practical Tips

### Do's

✅ [Tip 1]  
✅ [Tip 2]  
✅ [Tip 3]

### Don'ts

❌ [Pitfall 1]  
❌ [Pitfall 2]  
❌ [Pitfall 3]

# Key Takeaways {.tabset .tabset-fade .tabset-pills}

## Summary

### Main Points

1. **[Topic]**: [Key point summary]
2. **[Topic]**: [Key point summary]
3. **[Topic]**: [Key point summary]

### Critical Success Factors

- [Factor 1]
- [Factor 2]
- [Factor 3]

## Quick Reference

### Design Checklist

```{r checklist}
checklist <- data.frame(
  Phase = c("Phase 1", "Phase 2", "Phase 3", "Phase 4"),
  Key_Check = c(
    "Check 1",
    "Check 2",
    "Check 3",
    "Check 4"
  ),
  Typical_Value = c(
    "Value/Range",
    "Value/Range",
    "Value/Range",
    "Value/Range"
  )
)

kable(checklist, caption = "Design Checklist") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

### Common Pitfalls

| Pitfall | Symptom | Solution |
|---------|---------|----------|
| [Pitfall 1] | [How it manifests] | [How to fix] |
| [Pitfall 2] | [How it manifests] | [How to fix] |
| [Pitfall 3] | [How it manifests] | [How to fix] |

## For Interviews

### Key Concepts to Remember

1. [Concept 1 with brief explanation]
2. [Concept 2 with brief explanation]
3. [Concept 3 with brief explanation]

### Typical Interview Questions

**Q1**: [Question]  
**A**: [Concise answer with key points]

**Q2**: [Question]  
**A**: [Concise answer with key points]

**Q3**: [Question]  
**A**: [Concise answer with key points]

# References

<div id="refs"></div>

---

# Appendix {.tabset .tabset-fade .tabset-pills}

## Detailed Calculations

[Detailed mathematical derivations if needed]

## Additional Data

```{r additional-data}
# Additional supporting data or plots
```

## Source Code

### Utility Functions Used

```{r show-code, echo=TRUE, eval=FALSE}
# Show key code used in this chapter
# Available in shared_resources/scripts/
```

## Further Reading

### Recommended Resources

**Books:**
- [Book 1]
- [Book 2]

**Papers:**
- [Paper 1]
- [Paper 2]

**Standards:**
- [Standard 1]
- [Standard 2]

---

*Chapter last updated: `r Sys.Date()`*  
*Part of PA Design Reference Manual v1.0*
