---
title: "Spectrum Bands Test"
output: html_document
---

```{r setup, include=FALSE}
library(plotly)
```

## Standalone Spectrum Test

This is a minimal test to verify spectrum band visualization.

```{r spectrum_test, echo=TRUE}
# Create frequency bands
freq_bands <- list(
  list(name = "VHF", xmin = 0.03, xmax = 0.3, color = 'rgba(255,182,193,0.3)',
       label_x = log10(sqrt(0.03 * 0.3)), wavelength = "10-1 m", full_freq = "30-300 MHz",
       apps = "FM radio, Weather radio", icon = "📻"),
  list(name = "UHF", xmin = 0.3, xmax = 3, color = 'rgba(173,216,230,0.3)',
       label_x = log10(sqrt(0.3 * 3)), wavelength = "1-0.1 m", full_freq = "0.3-3 GHz",
       apps = "TV, GPS, Mobile phones", icon = "📱"),
  list(name = "SHF", xmin = 3, xmax = 30, color = 'rgba(144,238,144,0.3)',
       label_x = log10(sqrt(3 * 30)), wavelength = "100-10 mm", full_freq = "3-30 GHz",
       apps = "Satellite comms, Radar", icon = "🛰️"),
  list(name = "EHF", xmin = 30, xmax = 300, color = 'rgba(255,218,185,0.3)',
       label_x = log10(sqrt(30 * 300)), wavelength = "10-1 mm", full_freq = "30-300 GHz",
       apps = "Advanced radar, Experimental, 5G mmWave", icon = "📡")
)

# Create base plot
fig <- plot_ly() %>%
  layout(
    title = "Spectrum Bands Test - ALL VISIBLE by default",
    xaxis = list(
      title = "Frequency (GHz)",
      type = "log",
      range = c(log10(0.01), log10(300))
    ),
    yaxis = list(
      title = "Gain (dB)",
      range = c(-5, 40)
    ),
    showlegend = TRUE,
    hovermode = 'closest'
  )

# Add spectrum bands - VISIBLE by default for testing
band_idx <- 0
for(band in freq_bands) {
  band_idx <- band_idx + 1
  legend_name <- if(band_idx == 1) {
    sprintf('━━ Spectrum ━━<br>%s %s', band$icon, band$name)
  } else {
    sprintf('%s %s', band$icon, band$name)
  }
  
  # Build hover text for display (use line-breaks for clarity)
  hover_text <- sprintf(
    "%s %s (%s):\n%s\nWavelength: %s",
    band$icon, band$name, band$full_freq, band$apps, band$wavelength
  )

  # Add filled background that handles hover via 'fills'
  fig <- fig %>%
    add_trace(
      x = c(band$xmin, band$xmax, band$xmax, band$xmin, band$xmin),
      y = c(-5, -5, 40, 40, -5),
      type = 'scatter',
      mode = 'lines',
      fill = 'toself',
      fillcolor = band$color,
      line = list(width = 0),
      name = legend_name,
      legendgroup = 'spectrum',
      visible = TRUE,
      showlegend = (band_idx == 1),
      hoveron = 'fills',
      text = hover_text,
      hoverinfo = 'text',
      hoverlabel = list(font = list(size = 12))
    ) %>%
    add_trace(
      x = band$label_x,
      y = 42,
      type = 'scatter',
      mode = 'text',
      text = sprintf('<b>%s</b>', band$name),
      textfont = list(size = 12, color = 'rgba(0,0,0,0.95)', family = 'Arial Black'),
      textposition = 'top center',
      legendgroup = 'spectrum',
      visible = TRUE,
      showlegend = FALSE,
      hoverinfo = 'skip'
    )
}

# Add a simple reference line
fig <- fig %>%
  add_trace(
    x = c(0.01, 300),
    y = c(20, 20),
    type = 'scatter',
    mode = 'lines',
    line = list(color = 'black', width = 2),
    name = 'Reference',
    showlegend = TRUE
  )

fig
```

### Expected Results:

1. **All 4 spectrum bands should be VISIBLE** (VHF=pink, UHF=blue, SHF=green, EHF=orange)
2. **Band labels** (VHF, UHF, SHF, EHF) should appear at y=35
3. **Hover over any band** should show:
   - Large icon (16px)
   - Format: `[icon] Band Name (frequency range): applications | Wavelength: range`
   - Example: "📻 VHF (30-300 MHz): FM radio, Weather radio | Wavelength: 10-1 m"
4. **Click legend "━━ Spectrum ━━"** to toggle all bands on/off together

